---
title: "Partial RDA seed dispersers"
author: "Luis Antonio Arias Medellin"
date: "May 22, 2018"
output: html_document
---
##Partial RDA Seed dispersers
```{r,include=F,warning=F}
#Load libraries
library("xlsx")
library("rmarkdown")
library("ade4")
library("vegan")
library("gclus")
library("ape")
library("dplyr")
library("ggplot2")
library("tidyr")
library("gridExtra")
```
#This data base is using point-count sampling, in which an individual entered the forest and stopped 12 minutes to sample individulas (acoustic and visual identification), then continued walking and did the same procedure 3 times for small patches and 6 times for big patches. The description of each column can be see in the text file
```{r,include=F}
#Read file
base.df<-read.xlsx("Data bases/Frugivores_for_Luis.xlsx",sheetName=1)
```

#Prepare data base for analysis
```{r}
#I am going to aggregate the data by patch. I will obtain the mean of variables
seed_dispersers_variables<-aggregate(cbind(Area=base.df$Area,
                                           Prop_OldFor=base.df$Prop_OldFor,
                                           Pro_allFor=base.df$Pro_allFor,
                                           Alti=base.df$Alti)~
                     base.df$Site,FUN=mean)

#Rename columns
colnames(seed_dispersers_variables)[1]<-"Site"

#Extract explanatory variables and covariates, scaled and original variables
env_var<-data.frame(scale(seed_dispersers_variables[,c("Area","Prop_OldFor","Pro_allFor")])) 
raw.env.data<-data.frame(seed_dispersers_variables[,c("Area","Prop_OldFor","Pro_allFor")])
covariate<-data.frame(scale(seed_dispersers_variables[,"Alti"]))
raw.covariate<-data.frame(seed_dispersers_variables[,"Alti"])
colnames(raw.covariate)<-"elevation"


#I am going to aggregate species by patch.
dispersers.hell<-aggregate(cbind(Turdus_grayi=base.df$Turdus_grayi,
                                 Catharus_aurantiirostris=base.df$Catharus_aurantiirostris,
                                 Myadestes_melanops=base.df$Myadestes_melanops,
                                 Aulacorhynchus_prasinus=base.df$Aulacorhynchus_prasinus,
                                 Momotus_momota=base.df$Momotus_momota,
                                 Geotrygon_montana=base.df$Geotrygon_montana,
                                 Turdus_assimilis=base.df$Turdus_assimilis,
                                 Saltator_maximus=base.df$Saltator_maximus)~base.df$Site,FUN=sum)

#Make a copy of original data base to use for only Turdus grayi abundance (analysis below)
raw.abundances<-dispersers.hell
```


#Pairs plot
```{r}
#Plot pairs of abundance of each seed disperser with environmental variables
pairs(cbind(dispersers.hell,log(seed_dispersers_variables[,c("Area")]),seed_dispersers_variables[,c("Prop_OldFor","Pro_allFor","Alti")]),panel=panel.smooth) #ASK WHAT DOES HE MEANS WITH PROP.ALL.FOR (I WOULD THINK PRIMARY AND SECONDARY FOREST)

#Species abundance need to be transformed to be analyzed in the RDA. Here I am using the Hellinger transformation (STUDY WHY I USED THIS TRANSFORMATION. pAPERS IN FOLDER BIBLIOGRAPHY/MULTIVARIATE ANALYSIS. THERE I CN FIND THE NUMERICAL ECOLOGY WITH R AND OTHER PAPERS)
dispersers.hell<-decostand(dispersers.hell[-1],"hellinger") #This is if I aggregate species abundance by patch

#Other ways to do the RDA
#dispersers.hell<-decostand(dispersers.hell,"hellinger") #This is if I use raw data
#dispersers.hell<-decostand(dispersers.hell[-1],"pa") #This is if I want to run a RDA with presence/abscence
#dispersers.hell<-sqrt(dispersers.hell[-1]) #This is for square root transformation
```

#Run RDA
Conditioned = proportion of variance explanied by the covariate
Constrained = proportion of variance explained by the explanatory variables
Unconstrained = proportion of variance not explained by the model
Importance of components: for RDAx, it is the proportion of variance explained by the model and it has the same number of RDA as the explanatory variables used (In this case 3 variables). For PCAx is the proportion not explained by the model and it has the same number of PC as the number of response variables (In this case 8 species).

#Interpretation
Constrained variance is too low (0.07) and unconstrained variance is high (0.84), meaning that the explanatory variables are not explaining  too much in the abundance of species.
Rsquared adjusted is too low
```{r,fig.width=8, fig.height=8}
#Make RDA
dispersers.prda<-rda(dispersers.hell,env_var,covariate) 

#Summary of the analysis
summary(dispersers.prda)

#Unadjusted R^2 retrieved from the rda result
R2<-RsquareAdj(dispersers.prda)$r.squared
R2

#Adjusted R^2 retrieved from the rda object. THIS IS THE ONE A SHOULD LOOK AT SINCE IT IS ADJUSTED
R2adj<-RsquareAdj(dispersers.prda)$adj.r.squared
R2adj

```

#Plot RDA results
```{r}
#Triplot of the rda results
#Scaling 1: distance triplot for dispersers and seed dispersers with arrows of 
#different colors
#plot(dispersers.rda,scaling=1,main="Triplot RDA",type="points")
#sp.sc<-scores(dispersers.rda,choices=1:2,scaling=1,display="sp")
#arrows(0,0,sp.sc[,1],sp.sc[,2],length=0,lty=1,col="red")
#spp.scr <- scores(dispersers.rda, display = "species", scaling = 1) # y
#text(1.1*spp.scr[,1:2], rownames(spp.scr[,1]), 
#     col="black", xpd=T)
#legend(x=1.2,y=1,legend=paste(1:8,rownames(spp.scr[1:8,]),sep=" "),bty="n",cex=0.8)

#Scaling 2: distance triplot for dispersers and seed dispersers with arrows of 
#different colors

jpeg("Triplot RDA seed dispersers.jpg",quality=100,width=1000,height=700)
par(mar=c(6,6,3,3),xpd=F)
#plot(dispersers.prda,scaling=2,main="Triplot RDA seed dispersers (scaling 2)",type="points",cex=2,cex.axis=2,cex.lab=2,cex.main=2)
plot(dispersers.prda,scaling=2,type="points",cex=2,cex.axis=2,cex.lab=2,xlim=c(0,0),ylim=c(-1.5,0.3))
sp.sc<-scores(dispersers.prda,choices=1:2,scaling=2,display="sp")
arrows(0,0,sp.sc[,1],sp.sc[,2],length=0,lty=1,col="red",lwd=3)
spp.scr <- scores(dispersers.prda, display = "species", scaling = 2) # y
text(1.1*spp.scr[,1:2], rownames(spp.scr[,1]), 
     col="black", xpd=T,cex=1.5)
#legend(x=1,y=1.6,legend=paste(1:8,gsub(pattern="_",replacement=" ",x=rownames(spp.scr[1:8,])),sep=" "),bty="n",cex=1.5)
legend(x=-1.4,y=0,legend=paste(1:8,gsub(pattern="_",replacement=" ",x=rownames(spp.scr[1:8,])),sep=" "),bty="n",cex=1.3)

dev.off()
```

#Permutation for testing significance of the model
In this case, the model is not significant
```{r}
anova.cca(dispersers.prda)

extractAIC(dispersers.prda)
#anova(dispersers.prda,by="axis",step=1000)
```



##Regression of abundance of Turdus grayi only
```{r,echo=FALSE}
turdus.grayi<-data.frame(raw.abundances[,"Turdus_grayi"])
colnames(turdus.grayi)<-"Turdus_grayi"


panel.hist <- function(x, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(usr[1:2], 0, 1.5) )
    h <- hist(x, plot = FALSE)
    breaks <- h$breaks; nB <- length(breaks)
    y <- h$counts; y <- y/max(y)
    rect(breaks[-nB], 0, breaks[-1], y, col = "cyan", ...)
}


panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits = digits)[1]
    txt <- paste0(prefix, txt)
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}
```

```{r}

pairs(cbind(turdus.grayi,raw.env.data,raw.covariate),lower.panel=panel.cor,diag.panel=panel.hist,panel=panel.smooth,main="Raw data")

pairs(cbind(turdus.grayi,log(raw.env.data),raw.covariate),lower.panel=panel.cor,diag.panel=panel.hist,panel=panel.smooth,main="Raw data")

pairs(cbind(turdus.grayi,log(raw.env.data$Area),asin(sqrt(raw.env.data$Prop_OldFor)),raw.covariate),lower.panel=panel.cor,diag.panel=panel.hist,panel=panel.smooth,main="Raw data")


```

```{r}
#Check for collinearity of the data
library("usdm")
vif(data.frame(turdus.grayi,log(raw.env.data),raw.covariate))


#Do poisson regression
#turdus.mod1<-glm(turdus.grayi$Turdus_grayi~scale(log(raw.env.data$Area))+scale(log(raw.env.data$Prop_OldFor))+scale(raw.covariate$elevation),poisson)
turdus.mod1<-glm(turdus.grayi$Turdus_grayi~scale(raw.env.data$Area)+scale(raw.env.data$Prop_OldFor)+scale(raw.covariate$elevation),poisson)
summary(turdus.mod1)

#Quasipoisson regression due to overdispersion
#turdus.mod1<-glm(turdus.grayi$Turdus_grayi~scale(log(raw.env.data$Area))+scale(log(raw.env.data$Prop_OldFor))+scale(raw.covariate$elevation),quasipoisson)
turdus.mod1<-glm(turdus.grayi$Turdus_grayi~scale(raw.env.data$Area)+scale(raw.env.data$Prop_OldFor)+scale(raw.covariate$elevation),quasipoisson)
summary(turdus.mod1)

#Order the data to draw a line in the plot
ordered.data<-cbind(turdus.grayi,raw.env.data,raw.covariate)
ordered.data<-ordered.data[order(ordered.data$Area),]
```

```{r,fig.width=8, fig.height=8}
#Extract the formula to draw line on plot
formula.turdus<-exp(turdus.mod1$coefficients[1]+turdus.mod1$coefficients[2]*scale(ordered.data$Area))

#Plot PATCH SIZE
plot(scale(ordered.data$Area),ordered.data$Turdus_grayi,xlab="Patch size",ylab="Abundance",xaxt="n")
par(new=T)
#Plot estimated line
plot(scale(ordered.data$Area),formula.turdus,type="l",col="red",lwd=2,ylim=c(min(ordered.data$Turdus_grayi),max(ordered.data$Turdus_grayi)),axes=F,ylab="",xlab="")

#Put labels in axis. Scaling is done by the follwing formula: (original value-mean)/sd
#I want to have a certain range of values represented in the x axis, but I need to obtain the numbers of the scaled values so that I can put them in the x axis
original.values<-seq(from=100,to=max(round(ordered.data$Area,-2)),by=200)
scaled.values<-(original.values-mean(ordered.data$Area))/sd(ordered.data$Area)

#Put values in the x axis
axis(side=1,labels=original.values,at=scaled.values)

##PLOT PROPORTION OF FOREST
#Extract the formula to draw line on plot
formula.turdus<-exp(turdus.mod1$coefficients[1]+turdus.mod1$coefficients[3]*scale(ordered.data$Prop_OldFor))

#Plot data
plot(scale(ordered.data$Prop_OldFor),ordered.data$Turdus_grayi,xlab="Proportion of forest in a 1 km radius",ylab="Abundance",xaxt="n")
par(new=T)
#Plot estimated line
plot(scale(ordered.data$Prop_OldFor),formula.turdus,type="l",col="red",lwd=2,ylim=c(min(ordered.data$Turdus_grayi),max(ordered.data$Turdus_grayi)),axes=F,ylab="",xlab="")

#Put labels in axis. Scaling is done by the follwing formula: (original value-mean)/sd
#I want to have a certain range of values represented in the x axis, but I need to obtain the numbers of the scaled values so that I can put them in the x axis
original.values<-seq(from=0,to=max(round(ordered.data$Prop_OldFor,1)),by=0.2)
scaled.values<-(original.values-mean(ordered.data$Prop_OldFor))/sd(ordered.data$Prop_OldFor)

#Put values in the x axis
axis(side=1,labels=original.values,at=scaled.values)
```

#Turdus grayi abundance plot
```{r}
library("GGally")
cbind(turdus.grayi,raw.env.data,raw.covariate) %>%
  transmute(Turdus_grayi=log(Turdus_grayi+1),log.area=log(Area),Prop_OldFor,elevation) %>% 
  ggpairs(upper=list(continuous="points",wrap("smooth", method = "lm")),lower=list(continuous="cor"))
  
ggsave("Graphs/Turdus grayi vs variables.jpg")
```



##Species richness
```{r}
#Take out the column that indicates site id
richness<-raw.abundances[,-1]

#Add a 1 when an animal is detected
richness[richness>0] <-1

#Sum the number of species per site
richness$richness<-rowSums(richness)

#Do a data base that has the richness along with the environmental varibables
richness<-cbind(richness[,ncol(richness)],raw.env.data,raw.covariate)
colnames(richness)[1]<-"richness"

#Plot pairs
pairs(richness,lower.panel=panel.cor,diag.panel=panel.hist,panel=panel.smooth)

#Create a column of the logarithm of patch size
richness$logArea<-log(richness$Area)

#Plot pairs
pairs(richness[,c("richness","logArea","Prop_OldFor","Pro_allFor","elevation")],lower.panel=panel.cor,diag.panel=panel.hist,panel=panel.smooth)

#Scale variables
richness.scaled<-richness %>%
  mutate(logArea=scale(Area),
         Prop_OldFor=scale(Prop_OldFor),
         Pro_allFor=scale(Pro_allFor),
         elevation=scale(elevation)) %>% 
  dplyr::select(-Area)

formula.all<-richness~logArea+Prop_OldFor+Pro_allFor+elevation

#Make poisson regression including all the variables
summary(richnes.model1<-glm(formula.all,data=richness.scaled,family="poisson"))

#Make poisson regression exluding area
summary(richnes.model2<-update(richnes.model1, . ~ . -logArea))

#Make poisson regression excluding proportion of all forest
summary(richnes.model3<-update(richnes.model2, . ~ . -Pro_allFor))

#Make poisson regression excluding proportion of old growth forest. 
#I selected Poisson model since the resudal deviance is lower than the degrees of freedom; therefore, there is no overdispersion
summary(richnes.model4<-update(richnes.model3, . ~ . -Prop_OldFor))

#Check assumptions
par(mfrow=c(2,2))
plot(richnes.model4)

dev.off()

#Check normality of data
library("car")
qqPlot(richnes.model4$residuals)


#Select best random factor model using AIC
models.aic<-AIC(richnes.model1,richnes.model2,richnes.model3,richnes.model4)
library("MuMIn")
round(Weights(models.aic),digits=2) #Best model includes only proportion of forest 

rm(list = ls()[grep("^richnes.model", ls())])
```

# Plot of Species richness vs patch size and proportion of forest
```{r,fig.width=8, fig.height=8}
jpeg("Richness vs patch size and prop forest.jpg",quality=100,width=1200,height=700)

#Patch size
par(mar=c(6,6,3,3),xpd=F,mfrow=c(1,2))
plot(richness$logArea,richness$richness,xlab="Log patch size",ylab="Species richness",cex=2,cex.axis=2,cex.lab=2)

abline(glm(richness ~ logArea, data = richness,family="poisson"),col="red",lwd=2)

text(x=7,y=5,"a)",cex=2)

#Proportion of forest
par(mar=c(6,6,3,3),xpd=F)
plot(richness$Prop_OldFor,richness$richness,xlab="Proportion of old growth forest",ylab="Species richness",cex=2,cex.axis=2,cex.lab=2)

abline(glm(richness ~ Prop_OldFor, data = richness,family="poisson"),col="red",lwd=2)

text(x=0.8,y=5,"b)",cex=2)

dev.off()

p1<-richness %>% 
  ggplot(aes(x=logArea,y=richness)) +
  geom_point() + 
  geom_smooth(method="glm",method.args=list(family="poisson"),col="red") + 
  xlab("Log patch size") +
  ylab("Richness")

p2<-richness %>% 
  ggplot(aes(x=Pro_allFor,y=richness)) +
  geom_point() + 
  geom_smooth(method="glm",method.args=list(family="poisson"),col="red") + 
  xlab("Proportion of forest") +
  ylab("Richness")

p3<-grid.arrange(p1,p2,ncol=2)

ggsave("Richness vs patch size and prop forest.jpg",plot=p3)

rm(p1,p2,p3)
```


##Species abundance
```{r}
#Take out the column that indicates site id
library("reshape2")

#Create data frame with aggregated abundance of all species per patch
abundance<-melt(raw.abundances, id.vars="base.df$Site")
colnames(abundance)<-c("Site","Species","Abundance")
abundance<-aggregate(abundance$Abundance,by=list(Site=abundance$Site),FUN=sum)
colnames(abundance)[2]<-"Abundance"

abundance<-merge(abundance,seed_dispersers_variables,by="Site",all=F)
abundance$logArea<-log(abundance$Area)

#Plot pairs
pairs(abundance[,c(-1)],lower.panel=panel.cor,diag.panel=panel.hist,panel=panel.smooth)

```

#Poisson regression for abundance data
```{r}
#Scale variables
abundance.scaled<- abundance %>% 
  dplyr::select(Abundance,logArea,Pro_allFor,Prop_OldFor,Alti) %>% 
  mutate(logArea=scale(logArea),
         Pro_allFor=scale(Pro_allFor),
         Prop_OldFor=scale(Prop_OldFor),
         Alti=scale(Alti))

  #ggpairs(abundance)

#Model with all variables
#Check whether the model has over-dispersion or under-dispersion. If the Residual Deviance is greater than the degrees of freedom, then over-dispersion exists. https://newonlinecourses.science.psu.edu/stat504/node/57/
summary(mod.1<-glm(Abundance~logArea + Pro_allFor + Prop_OldFor + Alti,data=abundance.scaled,family=poisson(link="log")))

#I see there is over-dispersion of the data, so I will use a quasi-poisson model
summary(mod.1<-glm(Abundance~logArea + Pro_allFor + Prop_OldFor + Alti,data=abundance.scaled,family=quasipoisson(link="log")))

#Take out proportion of all forest
summary(mod.2<-update(mod.1,.~.-Pro_allFor))

#Take out elevation
summary(mod.3<-update(mod.2,.~.-Alti))

#Take out elevation
summary(mod.4<-update(mod.3,.~.-logArea))

#Quasipoisson models do not have AIC, so I will check assumptions of last model. I DO NOT KNOW HOW TO CHECK THE ASSUMPTIONS OF THE MODEL. 
plot(mod.4)
```


#Plot
```{r}
p1<-abundance %>% 
  ggplot(aes(x=logArea,y=Abundance)) +
  geom_point() + 
  geom_smooth(method="glm",method.args=list(family="quasipoisson"),col="red") + 
  xlab("Log patch size")

p2<-abundance %>% 
  ggplot(aes(x=Pro_allFor,y=Abundance)) +
  geom_point() + 
  geom_smooth(method="glm",method.args=list(family="quasipoisson"),col="red") + 
  xlab("Proportion of forest")

p3<-grid.arrange(p1,p2,ncol=2)

ggsave("Abundance vs patch size and prop forest.jpg",plot=p3)

rm(p1,p2,p3)

```

